---
title: "CalCOFI_results_20210715"
author: "Zack Gold"
date: "7/15/2021"
output: html_document
---
# Prep Code
## Load Libraries
```{r, hide = TRUE}


library(tidyverse)
library(rstanarm)
options(mc.cores = parallel::detectCores())
library(bayesplot)
library(modelr)
library(tidybayes)
library(wesanderson)
library(dendextend)
library(ggpmisc)
library(here)
library(vegan)
library(ggdendro)
library(rioja)
library(cluster)
library(ggrepel)
library(grid)
library(RColorBrewer)
library(ggallin)
library(patchwork)
library(ggpattern)
library(corrplot)
library(gclus)
library(lattice)
library(gplots)
```

## Color Palletes
```{r, Color Palletes ,results='hide'}
# Colors
wes_palette("Darjeeling2") -> dar2
wes_palette("Darjeeling1") -> dar1
wes_palette("FantasticFox1") -> moon1
wes_palette("Royal1") -> br2
wes_palette("Rushmore1") -> br1
wes_palette("IsleofDogs1") -> id1
wes_palette("Royal2") -> id2

dar3 <- c(dar2[2], dar1[2], dar2[4], dar2[3], moon1[5])
#show_col(dar3)

dar4 <- c(id1[1], dar1[2], dar2[2], dar2[3])
#show_col(dar4)
wes_palette("Zissou1") -> moon2
moon3 <- c(id1[1], moon2[1], br2[4], br1[4], br1[3], moon2[5])
#show_col(moon3)

col_27 <-
  c(dar1,
    dar2,
    id1,
    id2,
    moon2[1],
    moon1[5],
    br1[3],
    br2[1],
    moon2[3],
    "darkseagreen",
    "dodgerblue")
col_27[10] <- br2[4]
col_27[13] <- br1[4]

col_23 <- c(dar1, dar2, id1, id2, moon2[1], moon1[5])

palette.new <- read_csv(here("data", "alternative.pallete.csv"))
palette.new %>% 
mutate(
.,
color = case_when(
Type == "Benthic North" ~ "burlywood4",
Type == "Benthic Central"  ~ "burlywood3"  ,
Type == "Benthic Cosmopolitan"  ~
"burlywood1"  ,
Type == "Benthic South"  ~ "darksalmon"  ,
Type == "Mesopelagic North" ~ "#08519c",
Type == "Mesopelagic South"  ~ "#cb181d" ,
Type == "Mesopelagic Cosmopolitan"  ~ "#6baed6" ,
Type == "Mesopelagic Central" ~ "#3182bd" ,
Type == "Coastal Pelagic Central" ~ "#969696",
Type == "Coastal Pelagic North" ~ "black",
Type == "Cosmopolitan"   ~ "grey80"   ,
TRUE ~ "white"
)) -> palette.new
library(scales)
show_col(palette.new$color)

palette.new3 <-
  read_csv(here("data", "alternative.pallete.csv")) %>% filter (!str_detect(Type, "Coastal Pelagic North"))

palette.new3 %>% 
mutate(
.,
color = case_when(
Type == "Benthic North" ~ "burlywood4",
Type == "Benthic Central"  ~ "burlywood3"  ,
Type == "Benthic Cosmopolitan"  ~
"burlywood1"  ,
Type == "Benthic South"  ~ "darksalmon"  ,
Type == "Mesopelagic North" ~ "#08519c",
Type == "Mesopelagic South"  ~ "#cb181d" ,
Type == "Mesopelagic Cosmopolitan"  ~ "#6baed6" ,
Type == "Mesopelagic Central" ~ "#3182bd" ,
Type == "Coastal Pelagic Central" ~ "#969696",
Type == "Coastal Pelagic North" ~ "black",
Type == "Cosmopolitan"   ~ "grey80"   ,
TRUE ~ "white"
)) -> palette.new3

show_col(palette.new3$color)

pallete.fib1b <-
  palette.new %>% filter (!str_detect(Type, "Benthic Cosmopolitan")) %>% filter (!str_detect(Type, "Coastal Pelagic North"))


```


### SST Temp plot
```{r}



metadata2 <-
read.csv(file = "../data/calcofi_metadata_analysis_20210907.csv")


metadata2  %>%
mutate(
.,
station = fct_recode(
Sta_ID,
"Pt. Conception" =  "080.0 060.0" ,
"San Nicholas Island" = "086.7 050.0" ,
"San Diego Inshore" = "093.3 030.0",
"San Diego Offshore" = "093.3 060.0"
)
) %>%
ggplot(., aes(
x = Year,
y = two_month_sst,
group = station,
color = station
)) +
geom_line(size = 1) +
scale_color_manual(values = dar4) +
annotate(
"rect",
xmin = 2014,
xmax = 2019,
ymin = 11.5,
ymax = 19.5,
alpha = .2,
fill = "darkred"
) +
theme_bw() +
theme(
panel.background = element_rect(fill = 'white', colour = 'white'),
panel.grid.major = element_blank(),
panel.grid.minor = element_blank(),
axis.text.x = element_text(
size = 12,
angle = 30,
hjust = 1
),
axis.text.y = element_text(size = 12, angle = 0),
axis.title.x = element_text(size = 16,  face = "bold"),
axis.title.y = element_text(size = 16, angle = 90, face = "bold"),
strip.background = element_rect(fill = 'grey90'),
strip.text.x = element_text(size = 15, face = 'italic'),
legend.text = element_text(size = 12),
legend.position = c(0.5, 0.8),
legend.title = element_blank(),
legend.box = "vertical"
) + ylab(" SST ˚C")  -> temp_plot

temp_plot


```


## Trait Data
```{r}

#Trait Data
trait_edna_1 <-
read.csv("../data/habitat_association_to_check_art.csv", header = T)
as.data.frame(trait_edna_1) -> trait_edna

trait_edna %>%
arrange(ID_main) %>% mutate(name = str_replace(ID_main, " ", ".")) %>%  mutate(., ID_main = str_replace(ID_main, "AANOTHER", "sp."))   %>% dplyr::select(-Type) %>%
mutate(Type = str_c(Habitat, " ", Range)) -> trait_edna


trait_edna %>%
mutate(
.,
species_color = case_when(
Type == "Benthic North" ~ "burlywood4",
Type == "Benthic Central"  ~ "burlywood3"  ,
Type == "Benthic Cosmopolitan"  ~
"burlywood1"  ,
Type == "Benthic South"  ~ "darksalmon"  ,
Type == "Mesopelagic North" ~ "#08519c",
Type == "Mesopelagic South"  ~ "#cb181d" ,
Type == "Mesopelagic Cosmopolitan"  ~ "#6baed6" ,
Type == "Mesopelagic Central" ~ "#3182bd" ,
Type == "Coastal Pelagic Central" ~ "#969696",
Type == "Coastal Pelagic North" ~ "black",
Type == "Cosmopolitan"   ~ "grey80"   ,
TRUE ~ "white"
)
) -> trait_edna


```



### STAN Data

Code is commented out where it takes signficant computational power to run analyses. Outputs are saved in a google document for convenience. All code should run when uncommented.

```{r, function for transforming eDNA Index,results='hide'}
#load(here("data","Stan_Fit_combined_20210929_1338.RData"))

```

```{r, Format Stan Data for SST Comparisons, results='hide'}
#### SST Data
# 
# stanMod <- Output$stanMod
# 
# pars2 <- rstan::extract(stanMod, par = "b_main_grid")
# 
# Stations <- Output$Stations
# Stations %>%
#      separate(station_id, into = c("Year","Rpt_Line","Rpt_Sta"), remove = F, sep="_") %>%
#      unite(., "station" , c("Rpt_Line","Rpt_Sta"), remove=FALSE) -> Stations
# 
#  species_mapping <-  Output$species_mapping %>%  mutate(., ID_main = str_replace(ID_main, "AANOTHER","sp."))
# 
#  rm(Output)
# rm(stanMod)
# 
# pars2$b_main_grid %>%
#   as_tibble() %>%
#   pivot_longer(., cols = `1.1`:`84.56`, names_to="station.species", values_to = "est") %>%
#   separate(station.species, into = c("station_idx","sp_index_main"), remove = F, sep="\\.") %>%
#   mutate(., station_idx = as.numeric(station_idx),
#          sp_index_main = as.numeric(sp_index_main)) %>%
#    left_join(species_mapping) %>%
#      left_join(Stations) -> b_grid_tibble
# 
# 
# saveRDS(b_grid_tibble, here("data","b_grid_tibble_rpkn.RDS"))

#b_grid_tibble <- readRDS(file = here("data", "b_grid_tibble_rpkn.RDS"))

```

### Year - Species Mean
```{r}
# b_grid_tibble %>%
#   mutate(., log_est = log(est + 0.01)) %>%
#   group_by(Year, ID_main) %>%
#   dplyr::summarise(
#   mean = mean(est),
#   median = median(est),
#   sd = sd(est),
#   p05 = quantile(est, probs = 0.05),
#   p95 = quantile(est, probs = 0.95)
#   ) %>%
#   mutate(mean_log = log(mean + 0.1), sd_log = log(sd + 0.1)) -> b_ID_main_out
# 
# 
# saveRDS(b_ID_main_out, here("data","b_ID_main_out.RDS"))

b_ID_main_out <- readRDS(file = here("data", "b_ID_main_out.RDS"))
  

```

### Year - Type Mean
```{r}
# b_grid_tibble %>%
#   mutate(., log_est = log(est + 0.01)) %>%
#   left_join(trait_edna, by = c("ID_main")) %>%
#   group_by(Year, Type) %>%
#   dplyr::summarise(
#   mean = mean(est),
#   sum = sum(est),
#   mean_log = mean(log_est),
#   sd_log = sd(log_est),
#   p05_log = quantile(log_est, probs = 0.05),
#   p95_log = quantile(log_est, probs = 0.95)
#   ) -> b_ID_main_out_type
# 
# b_ID_main_out_type %>%
#   mutate(., log_sum = log(sum + 0.01)) -> b_ID_main_out_type
# 
# 
# saveRDS(b_ID_main_out_type, here("data","b_ID_main_out_type.RDS"))

b_ID_main_out_type <- readRDS(file = here("data", "b_ID_main_out_type.RDS"))

```

### Year - Site - Species
```{r, results='hide'}

# 
# b_grid_tibble %>%
# mutate(., log_est = log10(est + 0.01)) %>%
# group_by(Year, station, ID_main) %>%
# dplyr::summarise(
# mean = mean(est),
# median = median(est),
# sd = sd(est),
# p05 = quantile(est, probs = 0.05),
# p95 = quantile(est, probs = 0.95)
# ) %>%
# mutate(mean_log = log10(mean + 0.01), sd_log = log10(sd + 0.01)) -> b_ID_main_out_site
# 
# b_ID_main_out_site$Year  <- as.numeric(b_ID_main_out_site$Year)
# 
# saveRDS(b_ID_main_out_site, here("data","b_ID_main_out_site.RDS"))

b_ID_main_out_site <- readRDS(file = here("data", "b_ID_main_out_site.RDS"))

# Format Data
# test_nest_a <- b_ID_main_out_site %>%
# mutate(.,
# Normalized.biomass = mean,
# log_Normalized.biomass = mean_log) %>%
# left_join(two_month_sst) %>%
# mutate(PA_biomass = if_else(Normalized.biomass > 0.01, 1, 0)) %>%  #NOTE change to small value, rather than actual zero, given model output
# ungroup() %>%
# mutate(temp_std = (two_month_sst - mean(two_month_sst)) / sd(two_month_sst)) %>%
# mutate(., ID_main = str_replace(ID_main, "AANOTHER", "sp.")) %>%
# group_by(ID_main) %>%
# nest
# 
# saveRDS(test_nest_a, here("data","test_nest_a.RDS"))

test_nest_a <- readRDS(file = here("data", "test_nest_a.RDS"))
#Format Data
# test_nest_a_wc <- b_ID_main_out_site %>%
# mutate(.,
# Normalized.biomass = mean,
# log_Normalized.biomass = mean_log) %>%
# left_join(mean_100_T) %>%
# mutate(PA_biomass = if_else(Normalized.biomass > 0.01, 1, 0)) %>%  #NOTE change to small value, rather than actual zero, given model output
# ungroup() %>%
# mutate(temp_std = (mean_100_T - mean(mean_100_T)) / sd(mean_100_T)) %>%
# mutate(., ID_main = str_replace(ID_main, "AANOTHER", "sp.")) %>%
# group_by(ID_main) %>%
# nest
# 
# saveRDS(test_nest_a_wc, here("data","test_nest_a_wc.RDS"))



```

### MHW Comparison - Species
```{r, Format Stan Data for MHW Comparisons, results='hide'}


##### MHW Data
# 
# pars2$b_main_grid  %>%
# as_tibble() %>% mutate(id = row_number()) %>%
# pivot_longer(.,
# cols = `1.1`:`84.56`,
# names_to = "station.species",
# values_to = "est") %>%
# separate(
# station.species,
# into = c("station_idx", "sp_index_main"),
# remove = F,
# sep = "\\."
# ) %>%
# mutate(
# .,
# station_idx = as.numeric(station_idx),
# sp_index_main = as.numeric(sp_index_main)
# ) %>%
# left_join(species_mapping) %>%
# left_join(Stations) -> b_grid_tibble_2
# 
# saveRDS(b_grid_tibble_2, here("data", "b_grid_tibble_2_rpkn.RDS"))

#b_grid_tibble_2 <-readRDS(file = here("data", "b_grid_tibble_2_rpkn.RDS"))

# b_grid_tibble_2 %>%
# mutate(
# .,
# MHW = case_when(
# Year == "2014" ~ "MHW",
# Year == "2015" ~ "MHW",
# Year == "2016" ~ "MHW",
# Year == "2017" ~ "MHW",
# Year == "2018" ~ "MHW",
# Year == "2019" ~ "MHW",
# TRUE ~ "Before"
# )
# ) %>%
# mutate(., log_est = log(est + 0.01)) %>%
# group_by(ID_main, MHW, id) %>%
# dplyr::summarise(mean_b_est_log = mean(log_est),
# mean_b_est = mean(est)) -> b_ID_main_dist_out
# 
# b_ID_main_dist_out %>% pivot_wider(
# values_from = c(mean_b_est_log, mean_b_est),
# names_glue = "{MHW}_{.value}",
# names_from = MHW
# ) %>%
# mutate(., `MHW-Before` = MHW_mean_b_est - Before_mean_b_est) -> b_ID_main_dist_out
# 
# 
# saveRDS(b_ID_main_dist_out, here("data","b_ID_main_dist_out.RDS"))

b_ID_main_dist_out <- readRDS(file = here("data", "b_ID_main_dist_out.RDS"))

```

### MHW - Site
```{r}
# b_grid_tibble_2 %>%
#   mutate(
#   .,
#   MHW = case_when(
#   Year == "2014" ~ "MHW",
#   Year == "2015" ~ "MHW",
#   Year == "2016" ~ "MHW",
#   Year == "2017" ~ "MHW",
#   Year == "2018" ~ "MHW",
#   Year == "2019" ~ "MHW",
#   TRUE ~ "Before"
#   )
#   ) %>%
#   mutate(., log_est = log(est + 0.01)) %>%
#   group_by(ID_main, MHW, id, station) %>%
#   dplyr::summarise(mean_b_est = mean(log_est)) -> b_ID_main_dist_out_station
#   
# b_ID_main_dist_out_station %>% 
#     pivot_wider(values_from = mean_b_est, names_from =
#   MHW) %>%
#   mutate(., `MHW-Before` = MHW - Before) -> b_ID_main_dist_out_station
#   
# saveRDS(b_ID_main_dist_out_station, here("data","b_ID_main_dist_out_station.RDS"))

b_ID_main_dist_out_station <- readRDS(file = here("data", "b_ID_main_dist_out_station.RDS"))
```

# MHW Comparisons

### Averaged Across Sites

```{r}
b_ID_main_dist_out  -> b_MHW

b_MHW %>% 
  group_by(ID_main) %>% 
  dplyr::summarise(mean=mean(`MHW-Before`), median=median(`MHW-Before`),sd=sd(`MHW-Before`), p05= quantile(`MHW-Before`, probs=0.05), p95= quantile(`MHW-Before`, probs=0.95)) -> MHW_species_summary
  
MHW_species_summary %>% 
  mutate(., Sig_MHW = case_when(p05 > 0 & mean > 0 & p95 > 0 ~"+",
                                p05 < 0 & mean < 0 & p95 < 0 ~"-",
                                TRUE ~"")) %>% 
  filter(., Sig_MHW != "")-> MHW_species_summary_sig

MHW_species_summary_sig %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors


MHW_species_summary_sig %>%
  left_join(trait_edna) -> MHW_species_summary_sig

MHW_species_summary_sig %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw


species_color_mhw$Type <- as.character(species_color_mhw$Type)
species_color_mhw$Type <- factor(species_color_mhw$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

species_color_mhw %>% arrange(Type) -> species_color_mhw

MHW_species_summary_sig$Type <- as.character(MHW_species_summary_sig$Type)
MHW_species_summary_sig$Type <- factor(MHW_species_summary_sig$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))




MHW_species_summary_sig %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant")) %>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new3$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association")) + scale_y_continuous(trans = pseudolog10_trans, breaks = c(-40, 0,40,80,400 )) +
  theme(axis.text.x = element_text(size=12,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=16),
          legend.title=element_text(size=18),
          axis.line=element_line(colour='black'), 
        legend.position = c(0.9, 0.7) )   -> mhw_differential_abundance
mhw_differential_abundance

ggsave(mhw_differential_abundance, file =here::here("analysis","figures","mhw_differential_abundance.png"), width = 12, height = 8)

```

### All Sites Individually
```{r}

b_ID_main_dist_out_station %>% 
  group_by(ID_main, station) %>% 
  dplyr::summarise(mean=mean(`MHW-Before`), median=median(`MHW-Before`),sd=sd(`MHW-Before`), p05= quantile(`MHW-Before`, probs=0.05), p95= quantile(`MHW-Before`, probs=0.95))  %>% mutate(.,
station = fct_recode(station,
"Pt. Conception" =  "80_60" ,
"San Nicholas Island" = "86.7_50" ,
"San Diego Inshore" = "93.3_30",
"San Diego Offshore" = "93.3_60")
) -> MHW_species_summary_site
  
MHW_species_summary_site %>% 
  mutate(., Sig_MHW = case_when(p05 > 0 & mean > 0 & p95 > 0 ~"+",
                                p05 < 0 & mean < 0 & p95 < 0 ~"-",
                                TRUE ~"")) -> MHW_species_summary_sig_site_all


MHW_species_summary_sig_site_all %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_all


MHW_species_summary_sig_site_all %>%
  left_join(trait_edna) -> MHW_species_summary_sig_all

MHW_species_summary_sig_all %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_all


species_color_mhw_all$Type <- as.character(species_color_mhw_all$Type)
species_color_mhw_all$Type <- factor(species_color_mhw_all$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

species_color_mhw_all %>% arrange(Type) -> species_color_mhw_all

MHW_species_summary_sig_all$Type <- as.character(MHW_species_summary_sig_all$Type)

MHW_species_summary_sig_all$Type <- factor(MHW_species_summary_sig_all$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_all$Type %>% unique() -> all_type

palette.new %>% 
 filter(., Type %in% all_type) -> palette.new8

palette.new8$Type <- factor(palette.new8$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

palette.new8 %>% arrange(Type) -> palette.new8


MHW_species_summary_sig_all %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant"))%>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new8$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +geom_text(aes(label=Sig_MHW, y = mean + 0.3 * sign(mean)),size=8)+
  theme(axis.text.x = element_text(size=14,angle= 75, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=20, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=24, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.y=element_text(size=18,face="bold"),
          legend.text=element_text(size=14),
          legend.title=element_text(size=18,face="bold"),
          axis.line=element_line(colour='black'),
        panel.spacing = unit(2, "lines"))  + facet_grid(station~ .)   -> mhw_differential_abundance_all
mhw_differential_abundance_all

```

### SD Offshore

```{r}

MHW_species_summary_sig_site_all %>% 
  filter(., station =="San Diego Offshore") -> sig_sd_off_mhw


sig_sd_off_mhw %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_south


sig_sd_off_mhw %>%
  left_join(trait_edna) -> MHW_species_summary_sig_south

MHW_species_summary_sig_south %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_south


species_color_mhw_south$Type <- as.character(species_color_mhw_south$Type)
species_color_mhw_south$Type <- factor(species_color_mhw_south$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

species_color_mhw_south %>% arrange(Type) -> species_color_mhw_south

MHW_species_summary_sig_south$Type <- as.character(MHW_species_summary_sig_south$Type)

MHW_species_summary_sig_south$Type <- factor(MHW_species_summary_sig_south$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_south$Type %>% unique() -> south_type

palette.new %>% 
 filter(., Type %in% south_type) -> palette.new4

MHW_species_summary_sig_south %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant"))%>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +
  theme(axis.text.x = element_text(size=12,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=12),
          axis.line=element_line(colour='black'), 
        legend.position = c(.9, 0.85) )  +ylim(-6,4) -> mhw_differential_abundance_south
mhw_differential_abundance_south

ggsave(mhw_differential_abundance_south, file =here::here("analysis","figures","mhw_differential_abundance_sd_off.png"), width = 12, height = 8)


```
### SD Inshore

```{r}

MHW_species_summary_sig_site_all %>% 
  filter(., station =="San Diego Inshore") -> sig_sd_in_mhw

sig_sd_in_mhw %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_south


sig_sd_in_mhw %>%
  left_join(trait_edna) -> MHW_species_summary_sig_south

MHW_species_summary_sig_south %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_south


species_color_mhw_south$Type <- as.character(species_color_mhw_south$Type)
species_color_mhw_south$Type <- factor(species_color_mhw_south$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

species_color_mhw_south %>% arrange(Type) -> species_color_mhw_south

MHW_species_summary_sig_south$Type <- as.character(MHW_species_summary_sig_south$Type)

MHW_species_summary_sig_south$Type <- factor(MHW_species_summary_sig_south$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_south$Type %>% unique() -> south_type

palette.new %>% 
 filter(., Type %in% south_type) -> palette.new4


palette.new4$Type <- factor(palette.new4$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

palette.new4 %>% arrange(Type)  ->palette.new4


MHW_species_summary_sig_south %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant"))%>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +
  theme(axis.text.x = element_text(size=12,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=12),
          axis.line=element_line(colour='black'), 
        legend.position = c(.9, 0.85) )  +ylim(-6,4) -> mhw_differential_abundance_south
mhw_differential_abundance_south

ggsave(mhw_differential_abundance_south, file =here::here("analysis","figures","mhw_differential_abundance_sd_off.png"), width = 12, height = 8)



```

### Legend of Species w/ Colors
```{r}
library(ggpubr)

MHW_species_summary_sig_south %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant"))%>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Species Habitat Association"))  +
  theme(axis.text.x = element_text(size=12,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=12),
          axis.line=element_line(colour='black'), 
        legend.position = c(.5, 0.5) )  +ylim(-6,4) -> legend_holder
leg <- get_legend(legend_holder) 
leg 
# Convert to a ggplot and print
as_ggplot(leg) + plot_layout(tag_level = 'new') -> leggers 
as_ggplot(leg)  -> leggest

```

### SD Inshore

```{r}
MHW_species_summary_sig_site_all %>% 
  filter(., station =="San Diego Inshore") -> sig_sd_in_mhw

sig_sd_in_mhw %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_sd_in


sig_sd_in_mhw %>%
  left_join(trait_edna) -> MHW_species_summary_sig_sd_in

MHW_species_summary_sig_sd_in %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_sd_in


MHW_species_summary_sig_sd_in$Type <- as.character(MHW_species_summary_sig_sd_in$Type)
MHW_species_summary_sig_sd_in$Type <- factor(MHW_species_summary_sig_sd_in$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_sd_in %>% arrange(Type) -> MHW_species_summary_sig_sd_in

MHW_species_summary_sig_sd_in$Type <- as.character(MHW_species_summary_sig_sd_in$Type)

MHW_species_summary_sig_sd_in$Type <- factor(MHW_species_summary_sig_sd_in$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_sd_in$Type %>% unique() -> south_type

palette.new %>% 
 filter(., Type %in% south_type) -> palette.new4

MHW_species_summary_sig_sd_in %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant")) %>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +
  theme(axis.text.x = element_text(size=8,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=16),
          legend.title=element_text(size=18),
          axis.line=element_line(colour='black'), 
        legend.position = c(.9, 0.8) )   -> mhw_differential_abundance_sd_in
mhw_differential_abundance_sd_in


```


### Pt. Conception

```{r}
MHW_species_summary_sig_site_all %>% 
  filter(., station =="Pt. Conception") -> sig_pc_mhw

sig_pc_mhw %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_pc


sig_pc_mhw %>%
  left_join(trait_edna) -> MHW_species_summary_sig_pc

MHW_species_summary_sig_pc %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_pc


MHW_species_summary_sig_pc$Type <- as.character(MHW_species_summary_sig_pc$Type)
MHW_species_summary_sig_pc$Type <- factor(MHW_species_summary_sig_pc$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_pc %>% arrange(Type) -> MHW_species_summary_sig_pc

MHW_species_summary_sig_pc$Type <- as.character(MHW_species_summary_sig_pc$Type)

MHW_species_summary_sig_pc$Type <- factor(MHW_species_summary_sig_pc$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_pc$Type %>% unique() -> south_type

palette.new %>% 
 filter(., Type %in% south_type) -> palette.new4

MHW_species_summary_sig_pc %>% 
          mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant")) %>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +
  theme(axis.text.x = element_text(size=8,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=12),
          legend.title=element_text(size=14),
          axis.line=element_line(colour='black'), 
        legend.position = c(.9, 0.8) ) +ylim(-6,4)  -> mhw_differential_abundance_pc
mhw_differential_abundance_pc


```

### San Nicholas Island

```{r}
MHW_species_summary_sig_site_all %>% 
  filter(., station =="San Nicholas Island") -> sig_sni_mhw

sig_sni_mhw %>% 
     left_join(trait_edna, by=c("ID_main")) %>% 
  arrange(desc(mean)) %>%  dplyr::select(species_color) -> MHW_species_summary_sig_colors_sni


sig_sni_mhw %>%
  left_join(trait_edna) -> MHW_species_summary_sig_sni

MHW_species_summary_sig_sni %>% 
  dplyr::select(Type, species_color) %>%  unique()-> species_color_mhw_sni


MHW_species_summary_sig_sni$Type <- as.character(MHW_species_summary_sig_sni$Type)
MHW_species_summary_sig_sni$Type <- factor(MHW_species_summary_sig_sni$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_sni %>% arrange(Type) -> MHW_species_summary_sig_sni

MHW_species_summary_sig_sni$Type <- as.character(MHW_species_summary_sig_sni$Type)

MHW_species_summary_sig_sni$Type <- factor(MHW_species_summary_sig_sni$Type, levels=c("Benthic North" ,"Benthic Central","Benthic Cosmopolitan","Benthic South", "Coastal Pelagic North", "Coastal Pelagic Central","Mesopelagic North","Mesopelagic Central" , "Mesopelagic Cosmopolitan"  ,"Mesopelagic South"))

MHW_species_summary_sig_sni$Type %>% unique() -> south_type

palette.new %>% 
 filter(., Type %in% south_type) -> palette.new4

MHW_species_summary_sig_sni %>% 
         mutate(., ID_main=recode(ID_main,`Sebastes`="Sebastes sp.",
                               `Stenobrachius nannochir`="Stenobrachius leucopsarus cold variant")) %>% 
  ggplot(., aes(y= mean, x= reorder(ID_main, -mean), fill=Type)) +  geom_bar(position="stack", stat="identity") +
  scale_fill_manual( values = palette.new4$color) + ylab(expression(Delta*"Abundance")) + guides(fill=guide_legend(title="Habitat Association"))  +
  theme(axis.text.x = element_text(size=8,angle= 45, hjust= 1), 
        panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.y=element_text(size=16, angle=0), 
          axis.title.x=element_blank(), 
          axis.title.y=element_text(size=20, angle=90,face="bold"),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=16),
          legend.title=element_text(size=18),
          axis.line=element_line(colour='black'), 
        legend.position = c(.9, 0.8) )   -> mhw_differential_abundance_sni
mhw_differential_abundance_sni


```




# Figure 2 MHW
```{r}

mhw_differential_abundance_all 

ggsave(filename = here("analysis","figures","Figure_3_MHW_dif_all.png"),
       width=14,
       height = 16,
       dpi = 400,
      units = c("in"))



```

# Focal Species
## Anchovy and Sardine


## Anchovy and Sardine -Site
```{r}
b_ID_main_out_site %>% 
 subset(., ID_main %in% c("Engraulis mordax","Sardinops sagax")) %>% 
  mutate(.,Species= fct_recode(ID_main,
   "Northern Anchovy"= "Engraulis mordax" , "Pacific Sardine"= "Sardinops sagax" ), Year=as.numeric(Year)) %>% 
  mutate(., station= fct_recode(station,
   "Pt. Conception"=  "80_60" , "San Nicholas Island"= "86.7_50" , "San Diego Inshore"="93.3_30", "San Diego Offshore"= "93.3_60")) -> b_long_all_edit

b_long_all_edit %>% 
  mutate(., median_log = log(median+0.1),sd_log = log(sd+0.1),p05_log = log(p05+0.1),p95_log = log(p95+0.1) ) -> b_long_all_edit_site

ggplot(data=b_long_all_edit_site, aes(x=Year, group=Species, color=Species)) + 
  scale_color_manual(values=c("navy","darkred")) +
  geom_point(aes(y=`median_log`),size=2, na.rm = TRUE) +
  geom_line(aes(y=`median_log`),size=0.5, na.rm = TRUE) +
  geom_errorbar(aes(ymin = p05_log, ymax = p95_log), width = 0.1) +
  facet_wrap(station~., ncol=1) +theme_bw() +		## makes background white
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=10, angle=20, hjust = 0.65), 
          axis.text.y=element_text(size=12, angle=0), 
          axis.title.x=element_text(size=16, angle=0,face="bold"), 
          axis.title.y=element_text(size=16, angle=90,face="bold"),
          strip.background = element_rect(fill='grey90'), 
          strip.text.x=element_text(size=14, face='italic'),
          legend.text=element_text(size=12),
          legend.title = element_blank(),
         legend.box = "vertical") + ylab("log(Abundance)")+ annotate("rect", xmin = 2014, xmax = 2019, ymin = -3, ymax = 10, alpha = .2,fill = "darkred")-> biomass_site_chovy_sard
biomass_site_chovy_sard




```


```{r}

metadata2  %>%
mutate(
.,
station = fct_recode(
Sta_ID,
"Pt. Conception" =  "080.0 060.0" ,
"San Nicholas Island" = "086.7 050.0" ,
"San Diego Inshore" = "093.3 030.0",
"San Diego Offshore" = "093.3 060.0"
)
) -> temp_holder

temp_holder %>% 
filter(., station == "Pt. Conception") -> conc_temp
temp_holder %>% 
filter(., station == "San Nicholas Island") -> sni_temp
temp_holder %>% 
filter(., station == "San Diego Inshore") -> sdi_temp
temp_holder %>% 
filter(., station == "San Diego Offshore") -> sdo_temp

```
## Anchovy Sardine Pt. Conception

```{r}
b_long_all_edit_site %>% 
  filter(., station ==  "Pt. Conception") -> b_long_all_edit_site_conception

coeff1 <- 2.4
coeff2 <- -24

biomass_conc <- ggplot(data=b_long_all_edit_site_conception, aes(x=Year)) + 
    geom_rect_pattern(
    aes(
      xmin=2014, ymin=-3, xmax=2019, ymax=20,
      pattern_angle   = I(45),
      pattern_colour  = I("grey90"),
      pattern_spacing = I(0.05),
      pattern_size    = I(0.25)
    ),
    pattern         = 'stripe',
    fill            = 'white',
    colour          = 'white',
    pattern_density = 0.3
  ) +
  scale_color_manual(values=c("navy","darkred")) +
  geom_point(aes(y=`median_log`, group=Species, color=Species),size=2, na.rm = TRUE) +
  geom_line(aes(y=`median_log`,group=Species, color=Species),size=1, na.rm = TRUE) +
  geom_errorbar(aes(ymin = p05_log, ymax = p95_log,group=Species, color=Species), width = 0.1)  +
geom_line(data=conc_temp, aes(x=Year, y=(coeff2+two_month_sst*coeff1)), size = 2, color= "black") +
scale_y_continuous(
    
    # Features of the first axis
    name = "log(Abundance)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(.-coeff2)/coeff1, name="SST ˚C")
  ) +
  theme_bw() +		## makes background white
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=10, angle=20, hjust = 0.65), 
          axis.text.y=element_text(size=12, angle=0), 
          axis.title.x=element_text(size=16, angle=0,face="bold"), 
          axis.title.y=element_text(size=16, angle=90,face="bold"),
          strip.background = element_rect(fill='grey90'), 
          strip.text.x=element_text(size=14, face='italic'),
          legend.text=element_text(size=12),
          legend.position = c(0.3, 0.83),
          legend.title = element_blank(),
         legend.box = "vertical")+facet_wrap(station~., ncol=1) 

biomass_conc

```

## Anchovy Sardine San Nicholas Island

```{r}
b_long_all_edit_site %>% 
  filter(., station ==  "San Nicholas Island") -> b_long_all_edit_site_sni


biomass_sni <- ggplot(data=b_long_all_edit_site_sni, aes(x=Year)) + 
    geom_rect_pattern(
    aes(
      xmin=2014, ymin=-3, xmax=2019, ymax=20,
      pattern_angle   = I(45),
      pattern_colour  = I("grey90"),
      pattern_spacing = I(0.05),
      pattern_size    = I(0.25)
    ),
    pattern         = 'stripe',
    fill            = 'white',
    colour          = 'white',
    pattern_density = 0.3
  ) +
  scale_color_manual(values=c("navy","darkred")) +
  geom_point(aes(y=`median_log`, group=Species, color=Species),size=2, na.rm = TRUE) +
  geom_line(aes(y=`median_log`,group=Species, color=Species),size=1, na.rm = TRUE) +
  geom_errorbar(aes(ymin = p05_log, ymax = p95_log,group=Species, color=Species), width = 0.1)  +
geom_line(data=sni_temp, aes(x=Year, y=(coeff2+two_month_sst*coeff1)), size = 2, color= "black") +
scale_y_continuous(
    
    # Features of the first axis
    name = "log(Abundance)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(.-coeff2)/coeff1, name="SST ˚C")
  ) +
  theme_bw() +		## makes background white
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=10, angle=20, hjust = 0.65), 
          axis.text.y=element_text(size=12, angle=0), 
          axis.title.x=element_text(size=16, angle=0,face="bold"), 
          axis.title.y=element_text(size=16, angle=90,face="bold"),
          strip.background = element_rect(fill='grey90'), 
          strip.text.x=element_text(size=14, face='italic'),
          legend.text=element_text(size=12),
          legend.position = c(0.3, 0.83),
          legend.title = element_blank(),
         legend.box = "vertical") +facet_wrap(station~., ncol=1) 

biomass_sni


```

## Anchovy Sardine San Diego Inshore

```{r}
b_long_all_edit_site %>% 
  filter(., station ==  "San Diego Inshore") -> b_long_all_edit_site_sdi


coeff1i <- 2.4
coeff2i <- -27


biomass_sdi <- ggplot(data=b_long_all_edit_site_sdi, aes(x=Year)) + 
    geom_rect_pattern(
    aes(
      xmin=2014, ymin=-3, xmax=2019, ymax=20,
      pattern_angle   = I(45),
      pattern_colour  = I("grey90"),
      pattern_spacing = I(0.05),
      pattern_size    = I(0.25)
    ),
    pattern         = 'stripe',
    fill            = 'white',
    colour          = 'white',
    pattern_density = 0.3
  ) +
  scale_color_manual(values=c("navy","darkred")) +
  geom_point(aes(y=`median_log`, group=Species, color=Species),size=2, na.rm = TRUE) +
  geom_line(aes(y=`median_log`,group=Species, color=Species),size=1, na.rm = TRUE) +
  geom_errorbar(aes(ymin = p05_log, ymax = p95_log,group=Species, color=Species), width = 0.1)  +
geom_line(data=sdi_temp, aes(x=Year, y=(coeff2i+two_month_sst*coeff1i)), size = 2, color= "black") +
scale_y_continuous(
    
    # Features of the first axis
    name = "log(Abundance)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(.-coeff2)/coeff1, name="SST ˚C")
  ) +
  theme_bw() +		## makes background white
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=10, angle=20, hjust = 0.65), 
          axis.text.y=element_text(size=12, angle=0), 
          axis.title.x=element_text(size=16, angle=0,face="bold"), 
          axis.title.y=element_text(size=16, angle=90,face="bold"),
          strip.background = element_rect(fill='grey90'), 
          strip.text.x=element_text(size=14, face='italic'),
          legend.text=element_text(size=12),
          legend.position = c(0.3, 0.83),
          legend.title = element_blank(),
         legend.box = "vertical") + facet_wrap(station~., ncol=1) 

biomass_sdi



```

## Anchovy Sardine San Diego Offshore

```{r}
b_long_all_edit_site %>% 
  filter(., station ==  "San Diego Offshore") -> b_long_all_edit_site_sdo

biomass_sdo <- ggplot(data=b_long_all_edit_site_sdo, aes(x=Year)) + 
    geom_rect_pattern(
    aes(
      xmin=2014, ymin=-3, xmax=2019, ymax=20,
      pattern_angle   = I(45),
      pattern_colour  = I("grey90"),
      pattern_spacing = I(0.05),
      pattern_size    = I(0.25)
    ),
    pattern         = 'stripe',
    fill            = 'white',
    colour          = 'white',
    pattern_density = 0.3
  ) +
  scale_color_manual(values=c("navy","darkred")) +
  geom_point(aes(y=`median_log`, group=Species, color=Species),size=2, na.rm = TRUE) +
  geom_line(aes(y=`median_log`,group=Species, color=Species),size=1, na.rm = TRUE) +
  geom_errorbar(aes(ymin = p05_log, ymax = p95_log,group=Species, color=Species), width = 0.1)  +
geom_line(data=sdo_temp, aes(x=Year, y=(coeff2+two_month_sst*coeff1)), size = 2, color= "black") +
scale_y_continuous(
    
    # Features of the first axis
    name = "log(Abundance)",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(~(.-coeff2)/coeff1, name="SST ˚C")
  ) +
  theme_bw() +		## makes background white
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=10, angle=20, hjust = 0.65), 
          axis.text.y=element_text(size=12, angle=0), 
          axis.title.x=element_text(size=16, angle=0,face="bold"), 
          axis.title.y=element_text(size=16, angle=90,face="bold"),
          strip.background = element_rect(fill='grey90'), 
          strip.text.x=element_text(size=14, face='italic'),
          legend.text=element_text(size=12),
          legend.position = c(0.3, 0.83),
          legend.title = element_blank(),
         legend.box = "vertical") + facet_wrap(station~., ncol=1) 

biomass_sdo

```
# Figure 3
```{r}


biomass_conc+biomass_sni+biomass_sdi+biomass_sdo +plot_layout(ncol=1)

ggsave(file =here::here("analysis","figures","Figure_4_anchovy_sardine_time_combined.png"), width = 8, height = 16)

```








# All Sites Combined NMDS & Heatmap

### Pre processing NMDs and Heatmaps
```{r, functions used, include=FALSE}
#Another version of the dendrogram.  

dendro_data_k <- function(hc, k) {
  
  hcdata    <-  ggdendro::dendro_data(hc, type = "rectangle")
  seg       <-  hcdata$segments
  labclust  <-  cutree(hc, k)[hc$order]
  segclust  <-  rep(0L, nrow(seg))
  heights   <-  sort(hc$height, decreasing = TRUE)
  height    <-  mean(c(heights[k], heights[k - 1L]), na.rm = TRUE)
  
  for (i in 1:k) {
    xi      <-  hcdata$labels$x[labclust == i]
    idx1    <-  seg$x    >= min(xi) & seg$x    <= max(xi)
    idx2    <-  seg$xend >= min(xi) & seg$xend <= max(xi)
    idx3    <-  seg$yend < height
    idx     <-  idx1 & idx2 & idx3
    segclust[idx] <- i
  }
  
  idx                    <-  which(segclust == 0L)
  segclust[idx]          <-  segclust[idx + 1L]
  hcdata$segments$clust  <-  segclust
  hcdata$segments$line   <-  as.integer(segclust < 1L)
  hcdata$labels$clust    <-  labclust
  
  hcdata
}

set_labels_params <- function(nbLabels,
                              direction = c("tb", "bt", "lr", "rl"),
                              fan       = FALSE) {
  if (fan) {
    angle       <-  360 / nbLabels * 1:nbLabels + 90
    idx         <-  angle >= 90 & angle <= 270
    angle[idx]  <-  angle[idx] + 180
    hjust       <-  rep(0, nbLabels)
    hjust[idx]  <-  1
  } else {
    angle       <-  rep(0, nbLabels)
    hjust       <-  0
    if (direction %in% c("tb", "bt")) { angle <- angle + 45 }
    if (direction %in% c("tb", "rl")) { hjust <- 1 }
  }
  list(angle = angle, hjust = hjust, vjust = 0.5)
}

plot_ggdendro <- function(hcdata,
                          direction   = c("lr", "rl", "tb", "bt"),
                          fan         = FALSE,
                          scale.color = NULL,
                          branch.size = 1,
                          label.size  = 3,
                          nudge.label = 0.01,
                          expand.y    = 0.1) {
  
  direction <- match.arg(direction) # if fan = FALSE
  ybreaks   <- pretty(segment(hcdata)$y, n = 5)
  ymax      <- max(segment(hcdata)$y)
  
  ## branches
  p <- ggplot() +
    geom_segment(data         =  segment(hcdata),
                 aes(x        =  x,
                     y        =  y,
                     xend     =  xend,
                     yend     =  yend,
                     linetype =  factor(line),
                     colour   =  factor(clust)),
                 lineend      =  "round",
                 show.legend  =  FALSE,
                 size         =  branch.size)
  
  ## orientation
  if (fan) {
    p <- p +
      coord_polar(direction = -1) +
      scale_x_continuous(breaks = NULL,
                         limits = c(0, nrow(ggdendro::label(hcdata)))) +
      scale_y_reverse(breaks = ybreaks)
  } else {
    p <- p + scale_x_continuous(breaks = NULL)
    if (direction %in% c("rl", "lr")) {
      p <- p + coord_flip()
    }
    if (direction %in% c("bt", "lr")) {
      p <- p + scale_y_reverse(breaks = ybreaks)
    } else {
      p <- p + scale_y_continuous(breaks = ybreaks)
      nudge.label <- -(nudge.label)
    }
  }
  
  # labels
  labelParams <- set_labels_params(nrow(hcdata$labels), direction, fan)
  hcdata$labels$angle <- labelParams$angle
  
  p <- p +
    geom_text(data        =  ggdendro::label(hcdata),
              aes(x       =  x,
                  y       =  y,
                  label   =  label,
                  colour  =  factor(clust),
                  angle   =  angle),
              vjust       =  labelParams$vjust,
              hjust       =  labelParams$hjust,
              nudge_y     =  ymax * nudge.label,
              size        =  label.size,
              show.legend =  FALSE)
  
  # colors and limits
  if (!is.null(scale.color)) {
    p <- p + scale_color_manual(values = scale.color)
  }
  
  ylim <- -round(ymax * expand.y, 1)
  p    <- p + expand_limits(y = ylim)
  
  p
}


```
##### Chronological Clustering

<br />
```{r}
#Format for dend and clustering
b_ID_main_out %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., Normalized.biomass >0) %>% 
  ungroup() %>% 
  group_by(ID_main) %>% 
  dplyr::count() %>% 
  filter(., n> 5) -> species_to_keep

b_ID_main_out %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  #filter(., ID_main %in% species_to_keep$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide

eDNA_long_all_wide$Year -> rownames(eDNA_long_all_wide)
eDNA_long_all_wide[,-1] -> eDNA_long_all_wide

#Create a bray-curtis distance matrix
diss <- vegdist(eDNA_long_all_wide, method='bray')

#Run the chronological cluster; chclust is the main function from rioja
clust <- chclust(diss)


```

```{r}
##Look at the number of clusters.  Then make a category for each cluster
ddata1 <- dendro_data_k(clust,8)

labs <- ggdendro::label(ddata1)
moon4 <- c("black","navy",moon3,"darkred")
p1 <- plot_ggdendro(ddata1,
                    direction   = "tb",
                    label.size  = 2.5,
                    scale.color = moon4,
                    branch.size = 0.5,
                    expand.y    = 0.2)

p1 <- p1 + theme_void() + expand_limits(x = c(-1, 32))
p1

```
```{r, results='hide'}

#Recreate above chronological clustering dendrogram but using same format as species clustering for figure

BPA_9 <- data.frame(eDNA_long_all_wide)

dend2 <- BPA_9 %>% vegdist(., method='bray') %>% chclust() %>% as.dendrogram %>% 
  color_branches(k=6, col = moon3)

```

```{r}
#Color code names of species based on trait information
as.data.frame(moon3) -> mooner
mooner$clust <- seq(1,6)
labs %>% 
  left_join(mooner) -> labs

```

<br />

##### Species Clustering

<br />
```{r,fig.height=8, fig.width=8}
## Input is the distance matrix (abundance, etc), with years on x and species on y. and species names as the rownames.
BPA <- data.frame(t(eDNA_long_all_wide))

colnames(BPA) <- rownames(eDNA_long_all_wide)
#BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram 

BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram %>% sort(type = "nodes")

#Decide on the number of branches for species, k=3
dend <- BPA %>% vegdist(., method='bray')%>% hclust(method = "complete") %>% as.dendrogram %>% sort(type = "nodes") %>% 
  color_branches(k=5, col=dar3)
plot(dend)
```



```{r}
#Color code names of species based on trait information
as.data.frame(rownames(BPA)) -> species_names

species_names %>% 
  left_join(trait_edna, by=c("rownames(BPA)"="ID_main")) -> species_names

```

##### Figure S3 Heatmap

```{r, fig.height=16, fig.width=8}
#Color palette
some_col_func <- colorspace::sequential_hcl(100,palette="BluYl")


BPA2 <- as.matrix(log10(BPA))
BPA2[BPA2 < 0] <- NA



colseperator <- labs %>% 
  group_by(clust) %>% 
  dplyr::summarise(label =max(label)) %>% 
  left_join(labs)

jpeg(file = here("analysis","figures","S3_All_sites_heatmap.png")
     , width=16,
       height = 12,
       res = 400,
      units = c("in"))
gplots::heatmap.2(BPA2, 
                  main = "",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
dev.off()


gplots::heatmap.2(BPA2, 
                  main = "eDNA All Sites",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
```


##### Figure S4 NMDS

```{r}

b_ID_main_out %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
 # filter(., ID_main %in% species_to_keep$ID_main) %>% 
  #mutate(., ID_main = str_replace(ID_main," ",".")) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2

eDNA_long_all_wide2$Year -> rownames(eDNA_long_all_wide2)
eDNA_long_all_wide2[,-1] -> eDNA_long_all_wide2

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1 <- eDNA_long_all_wide2

mds <- vegan::metaMDS(sptran1,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)
```




```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide,scor)
score1$rowname <- rownames(score1)
labs$label -> years

as.data.frame(years) -> nmdscov
rownames(nmdscov) <- nmdscov$years
nmdscov$rowname <- rownames(nmdscov)
score2 <- merge(score1, nmdscov, by="rowname", all.x=T, all.y=F)
score2 <- score2[order(score2$year),] 

### from chron cluster
### assign a cluster to each year
score2$cluster <- as.character(labs$clust)


spdf1 %>% 
  dplyr::select(Type,species_color)%>% distinct() %>% 
  arrange(Type) %>%  dplyr::select(species_color) -> color_order_nmds
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path (data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2)) +
  geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= years, colour=cluster), size=10) + 
  scale_colour_manual(values = c("navy","darkred",moon3,color_order_nmds$species_color)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S4_All_sites_NMDS.png"),
       width=30,
       height = 30,
       dpi = 400,
      units = c("in"))
```

# All Sites Separate NMDS


```{r}

b_ID_main_out_site %>% 
  mutate(.,
station = fct_recode(station,
"Pt.C" =  "80_60" ,
"SaNI" = "86.7_50" ,
"SDIn" = "93.3_30",
"SDOf" = "93.3_60")
) %>% 
  dplyr::select(Year,  station, ID_main,Normalized.biomass=  mean) %>% 
 # filter(., ID_main %in% species_to_keep$ID_main) %>% 
  #mutate(., ID_main = str_replace(ID_main," ",".")) %>% 
  unite(., "Year", c("Year","station"), sep=":") %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2_sep

eDNA_long_all_wide2_sep %>% dplyr::select(Year) %>% as.data.frame()-> sampledf_year_site
eDNA_long_all_wide2_sep$Year -> rownames(sampledf_year_site)

sampledf_year_site %>% 
  separate(Year, into = c("Year","station"), sep=":", remove=FALSE) -> sampledf_year_site

sampledf_year_site %>% 
  mutate(., station2 = case_when(station =="Pt.C"~"80_60",
                                station =="SaNI"~"86.7_50",
                                station =="SDOf"~"93.3_60",
                                station =="SDIn"~"93.3_30"),
         Year=as.integer(Year)) %>%  
  left_join(two_month_sst, by=c("Year"="Year","station2"="station")) -> sampledf_year_site

eDNA_long_all_wide2_sep$Year -> rownames(eDNA_long_all_wide2_sep)
eDNA_long_all_wide2_sep[,-1] -> eDNA_long_all_wide2_sep

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1_sep <- eDNA_long_all_wide2_sep

mds <- vegan::metaMDS(sptran1_sep,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)

```

###### PERMANOVA

```{r}

adonis2(eDNA_long_all_wide2_sep~Year+two_month_sst+station, data=sampledf_year_site)

```

###### Homogeneity of Dispersions
```{r}

d_carn <- vegdist(eDNA_long_all_wide2_sep, method="bray") 
  
#Homogeneity of dispersions test
betadisper(d_carn, getElement(sampledf_year_site, "station"))
anova(betadisper(d_carn, getElement(sampledf_year_site, "station")))
broom::tidy(TukeyHSD(betadisper(d_carn, getElement(sampledf_year_site, "station")))) %>% 
  arrange(adj.p.value)
```


```{r}
#Homogeneity of dispersions test
betadisper(d_carn, getElement(sampledf_year_site, "Year"))
anova(betadisper(d_carn, getElement(sampledf_year_site, "Year")))
broom::tidy(TukeyHSD(betadisper(d_carn, getElement(sampledf_year_site, "Year")))) %>% 
  arrange(adj.p.value)

```


##### Figure S5 NMDS

```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide2_sep,scor)
score1$rowname <- rownames(score1)

score1 %>% 
separate(rowname, into=c("Year","Station"), sep=":", remove = F) -> score2

ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path(data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2, group=Station, colour=Station)) +
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= rowname, group=Station, colour=Station), size=10)+
    geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
    scale_colour_manual(values = c(color_order_nmds$species_color,dar4)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S5_All_sites_sep_NMDS.png"),
       width=50,
       height = 50,
       dpi = 400,
       limitsize = FALSE,
      units = c("in"))
```
### Figure Site Only
```{r}
score2 %>% 
  mutate(., station = case_when(Station =="Pt.C"~"Pt. Conception",
                                Station =="SaNI"~"San Nicholas Island",
                                Station =="SDOf"~"San Diego Offshore",
                                Station =="SDIn"~"San Diego Inshore")) -> score3

score3$station <- factor(score3$station, levels = c("Pt. Conception", "San Nicholas Island", "San Diego Inshore","San Diego Offshore"))
  
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path(data=score3, aes(x = NMDS1, y = NMDS2, group=station, colour=station),alpha=0.8) +
  geom_text_repel(data=score3, aes(x = NMDS1, y = NMDS2,label= Year, group=station, colour=station), size=10)+
    scale_colour_manual(values = c(dar4)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'))	+ guides(colour=guide_legend(title="Station"))

```

##### CAP ordination
<br />

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide2_sep ~ Year+two_month_sst+station , data=sampledf_year_site, dist="bray")
summary(vare.cap) -> cap_sum

#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide2_sep

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  top_n(10, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  separate(sample, into = c("Year","station"), sep=":", remove=FALSE) %>% 
  mutate(., Year =as.integer(Year)) %>% 
  left_join(sampledf_year_site) %>% 
   mutate(., Station = case_when(station =="Pt.C"~"Pt. Conception",
                                station =="SaNI"~"San Nicholas Island",
                                station =="SDOf"~"San Diego Offshore",
                                station =="SDIn"~"San Diego Inshore")) %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(shape=Station, col=Year), size =3) + scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2)) +xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) 

ggsave(filename = here("analysis","figures","CAP_ord_species.png"),
       width=24,
       height = 12,
       dpi = 400,
      units = c("in"))

```
##### CAP Environmental Variables

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide2_sep ~ Year+two_month_sst+station , data=sampledf_year_site, dist="bray")
summary(vare.cap) -> cap_sum

 
#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide2_sep

as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  mutate(., new_name= case_when(sample=="Year" ~"Year",
                                sample=="two_month_sst" ~"SST ˚C",
                                sample=="stationSaNI" ~ "San Nicholas Island",
                                sample =="stationSDIn" ~"San Diego Inshore",
                                sample=="stationSDOf"~"San Diego Offshore"  ))-> cap_vars

cap_vars <- as.data.frame(cap_vars)
rownames(cap_vars) <- cap_vars$new_name

# Now add the environmental variables as arrows
arrowmat <- cap_vars


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))
library(scales)

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  separate(sample, into = c("Year","station"), sep=":", remove=FALSE) %>% 
  mutate(., Year =as.integer(Year)) %>% 
  left_join(sampledf_year_site) %>% 
   mutate(., Station = case_when(station =="Pt.C"~"Pt. Conception",
                                station =="SaNI"~"San Nicholas Island",
                                station =="SDOf"~"San Diego Offshore",
                                station =="SDIn"~"San Diego Inshore")) %>% 
  ggplot(aes(x=CAP1, y=CAP2)) + geom_point(aes(shape=Station, col=Year), size =3) + scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))


ggsave(filename = here("analysis","figures","CAP_ord_sites.png"),
       width=16,
       height = 8,
       dpi = 400,
      units = c("in"))

```

# SD Offshore NMDS & Heatmap
##### Chronological Clustering

<br />
```{r}
#Format for dend and clustering
b_ID_main_out_site %>% 
mutate(.,
station = fct_recode(station,
"Pt. Conception" =  "80_60" ,
"San Nicholas Island" = "86.7_50" ,
"San Diego Inshore" = "93.3_30",
"San Diego Offshore" = "93.3_60")
) -> b_ID_main_out_site_f
  
b_ID_main_out_site_f %>%   
  filter(., station =="San Diego Offshore") %>% 
  dplyr::select(Year, ID_main,Normalized.biomass=  mean) %>% 
  filter(., Normalized.biomass >0) %>% 
  ungroup() %>% 
  group_by(ID_main) %>% 
  dplyr::count() %>% 
  filter(., n> 5) -> species_to_keep_sd

b_ID_main_out_site_f %>% 
    filter(., station =="San Diego Offshore") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sd$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide_sd

eDNA_long_all_wide_sd %>% dplyr::select(Year,station) %>% as.data.frame()-> sampledf_year_sd
eDNA_long_all_wide_sd$Year -> rownames(sampledf_year_sd)

sampledf_year_sd %>% 
  mutate(., station2 = case_when(station =="Pt. Conception"~"80_60",
                                station =="San Nicholas Island"~"86.7_50",
                                station =="San Diego Offshore"~"93.3_60",
                                station =="San Diego Inshore"~"93.3_30"),
         Year=as.integer(Year)) %>%  
  left_join(two_month_sst, by=c("Year"="Year","station2"="station")) -> sampledf_year_sd


eDNA_long_all_wide_sd$Year -> rownames(eDNA_long_all_wide_sd)
eDNA_long_all_wide_sd[,-1:-2] -> eDNA_long_all_wide_sd

#Create a bray-curtis distance matrix
diss_sd_off <- vegdist(eDNA_long_all_wide_sd, method='bray')

#Run the chronological cluster; chclust is the main function from rioja
clust_sd_off <- chclust(diss_sd_off)


```

```{r}
##Look at the number of clusters.  Then make a category for each cluster
ddata1 <- dendro_data_k(clust_sd_off,8)

labs <- ggdendro::label(ddata1)
moon4 <- c("black","navy",moon3,"darkred")
p1 <- plot_ggdendro(ddata1,
                    direction   = "tb",
                    label.size  = 2.5,
                    scale.color = moon4,
                    branch.size = 0.5,
                    expand.y    = 0.2)

p1 <- p1 + theme_void() + expand_limits(x = c(-1, 32))
p1

```
```{r, results='hide'}

#Recreate above chronological clustering dendrogram but using same format as species clustering for figure

BPA_9 <- data.frame(eDNA_long_all_wide_sd)

dend2 <- BPA_9 %>% vegdist(., method='bray') %>% chclust() %>% as.dendrogram %>% 
  color_branches(k=6, col = moon3)

```

```{r}
#Color code names of species based on trait information
as.data.frame(moon3) -> mooner
mooner$clust <- seq(1,6)
labs %>% 
  left_join(mooner) -> labs

```

<br />

##### Species Clustering

<br />
```{r,fig.height=8, fig.width=8}
## Input is the distance matrix (abundance, etc), with years on x and species on y. and species names as the rownames.
BPA <- data.frame(t(eDNA_long_all_wide_sd))

colnames(BPA) <- rownames(eDNA_long_all_wide_sd)
#BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram 

BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram %>% sort(type = "nodes")

#Decide on the number of branches for species, k=3
dend <- BPA %>% vegdist(., method='bray')%>% hclust(method = "complete") %>% as.dendrogram %>% sort(type = "nodes") %>% 
  color_branches(k=5, col=dar3)
plot(dend)
```



```{r}
#Color code names of species based on trait information
as.data.frame(rownames(BPA)) -> species_names

species_names %>% 
  left_join(trait_edna, by=c("rownames(BPA)"="ID_main")) -> species_names

```

##### Figure S6 Heatmap

```{r, fig.height=16, fig.width=8}
#Color palette
some_col_func <- colorspace::sequential_hcl(100,palette="BluYl")


BPA2 <- as.matrix(log10(BPA))
BPA2[BPA2 < 0] <- NA



colseperator <- labs %>% 
  group_by(clust) %>% 
  dplyr::summarise(label =max(label)) %>% 
  left_join(labs)

jpeg(file = here("analysis","figures","S6_SD_Offshore_heatmap.png")
     , width=16,
       height = 12,
       res = 400,
      units = c("in"))
gplots::heatmap.2(BPA2, 
                  main = "",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
dev.off()


gplots::heatmap.2(BPA2, 
                  main = "eDNA All Sites",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
```


##### Figure S7 NMDS

```{r}
b_ID_main_out_site_f %>% 
    filter(., station =="San Diego Offshore") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sd$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2_sd


eDNA_long_all_wide2_sd$Year -> rownames(eDNA_long_all_wide2_sd)
eDNA_long_all_wide2_sd[,-1:-2] -> eDNA_long_all_wide2_sd

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1 <- eDNA_long_all_wide2_sd

mds <- vegan::metaMDS(sptran1,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)

```




```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide2_sd,scor)
score1$rowname <- rownames(score1)
labs$label -> years

as.data.frame(years) -> nmdscov
rownames(nmdscov) <- nmdscov$years
nmdscov$rowname <- rownames(nmdscov)
score2 <- merge(score1, nmdscov, by="rowname", all.x=T, all.y=F)
score2 <- score2[order(score2$year),] 

### from chron cluster
### assign a cluster to each year
score2$cluster <- as.character(labs$clust)


spdf1 %>% 
  dplyr::select(Type,species_color)%>% distinct() %>% 
  arrange(Type) %>%  dplyr::select(species_color) -> color_order_nmds
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path (data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2)) +
  geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= years, colour=cluster), size=10) + 
  scale_colour_manual(values = c("navy","darkred",moon3,color_order_nmds$species_color)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S7_SD_offshore_NMDS.png"),
       width=30,
       height = 30,
       dpi = 400,
      units = c("in"))
```

##### CAP ordination
<br />

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide2_sd ~ two_month_sst + Year , data=sampledf_year_sd, dist="bray")
summary(vare.cap) -> cap_sum

#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide2_sd

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  top_n(5, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  mutate(., Station ="San Diego Offshore",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_sd) %>% 
  mutate(.,          `SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2))  + 
  geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
  scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2)) +xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) 

ggsave(filename = here("analysis","figures","CAP_ord_species_sdoff.png"),
       width=24,
       height = 12,
       dpi = 400,
      units = c("in"))

```
##### CAP Environmental Variables

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide2_sd ~ two_month_sst + Year , data=sampledf_year_sd, dist="bray")
summary(vare.cap) -> cap_sum

 
#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide2_sd

as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  mutate(., new_name= case_when(sample=="Year" ~"Year",
                                sample=="two_month_sst" ~"SST ˚C"))-> cap_vars

cap_vars <- as.data.frame(cap_vars)
rownames(cap_vars) <- cap_vars$new_name

# Now add the environmental variables as arrows
arrowmat <- cap_vars


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
 mutate(., Station ="San Diego Offshore",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_sd) %>% 
  mutate(.,`SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2)) +   geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
   scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))


ggsave(filename = here("analysis","figures","CAP_ord_sd_env.png"),
       width=16,
       height = 8,
       dpi = 400,
      units = c("in"))

```


# SD Inshore NMDS & Heatmap
##### Chronological Clustering

<br />
```{r}
b_ID_main_out_site_f %>%   
  filter(., station =="San Diego Inshore") %>% 
  dplyr::select(Year, ID_main,Normalized.biomass=  mean) %>% 
  filter(., Normalized.biomass >0) %>% 
  ungroup() %>% 
  group_by(ID_main) %>% 
  dplyr::count() %>% 
  filter(., n> 5) -> species_to_keep_sdi

b_ID_main_out_site_f %>% 
    filter(., station =="San Diego Inshore") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sdi$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide_sdi


eDNA_long_all_wide_sdi %>% dplyr::select(Year,station) %>% as.data.frame()-> sampledf_year_sdi
eDNA_long_all_wide_sdi$Year -> rownames(sampledf_year_sdi)

sampledf_year_sdi %>% 
  mutate(., station2 = case_when(station =="Pt. Conception"~"80_60",
                                station =="San Nicholas Island"~"86.7_50",
                                station =="San Diego Offshore"~"93.3_60",
                                station =="San Diego Inshore"~"93.3_30"),
         Year=as.integer(Year)) %>%  
  left_join(two_month_sst, by=c("Year"="Year","station2"="station")) -> sampledf_year_sdi

eDNA_long_all_wide_sdi$Year -> rownames(eDNA_long_all_wide_sdi)
eDNA_long_all_wide_sdi[,-1:-2] -> eDNA_long_all_wide_sdi

#Create a bray-curtis distance matrix
diss_sd_in <- vegdist(eDNA_long_all_wide_sdi, method='bray')

#Run the chronological cluster; chclust is the main function from rioja
clust_sd_in <- chclust(diss_sd_in)


```

```{r}
##Look at the number of clusters.  Then make a category for each cluster
ddata1 <- dendro_data_k(clust_sd_in,8)

labs <- ggdendro::label(ddata1)

moon4 <- c("black","navy",moon3,"darkred")
p1 <- plot_ggdendro(ddata1,
                    direction   = "tb",
                    label.size  = 2.5,
                    scale.color = moon4,
                    branch.size = 0.5,
                    expand.y    = 0.2)

p1 <- p1 + theme_void() + expand_limits(x = c(-1, 32))
p1

```
```{r, results='hide'}

#Recreate above chronological clustering dendrogram but using same format as species clustering for figure

BPA_9 <- data.frame(eDNA_long_all_wide_sdi)

dend2 <- BPA_9 %>% vegdist(., method='bray') %>% chclust() %>% as.dendrogram %>% 
  color_branches(k=6, col = moon3)

```

```{r}
#Color code names of species based on trait information
as.data.frame(moon3) -> mooner
mooner$clust <- seq(1,6)

labs %>% 
  left_join(mooner) -> labs

```

<br />

##### Species Clustering

<br />
```{r,fig.height=8, fig.width=8}
## Input is the distance matrix (abundance, etc), with years on x and species on y. and species names as the rownames.
BPA <- data.frame(t(eDNA_long_all_wide_sdi))

colnames(BPA) <- rownames(eDNA_long_all_wide_sdi)
#BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram 

BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram %>% sort(type = "nodes")

#Decide on the number of branches for species, k=3
dend <- BPA %>% vegdist(., method='bray')%>% hclust(method = "complete") %>% as.dendrogram %>% sort(type = "nodes") %>% 
  color_branches(k=5, col=dar3)
plot(dend)
```



```{r}
#Color code names of species based on trait information
as.data.frame(rownames(BPA)) -> species_names

species_names %>% 
  left_join(trait_edna, by=c("rownames(BPA)"="ID_main")) -> species_names

```

##### Figure S8 Heatmap

```{r, fig.height=16, fig.width=8}
#Color palette
some_col_func <- colorspace::sequential_hcl(100,palette="BluYl")


BPA2 <- as.matrix(log10(BPA))
BPA2[BPA2 < 0] <- NA



colseperator <- labs %>% 
  group_by(clust) %>% 
  dplyr::summarise(label =max(label)) %>% 
  left_join(labs)

jpeg(file = here("analysis","figures","S8_SD_Inshore_heatmap.png")
     , width=16,
       height = 12,
       res = 400,
      units = c("in"))
gplots::heatmap.2(BPA2, 
                  main = "",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
dev.off()


gplots::heatmap.2(BPA2, 
                  main = "eDNA SD Inshore",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
```


##### Figure S9 NMDS

```{r}
b_ID_main_out_site_f %>% 
    filter(., station =="San Diego Inshore") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sd$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2_sdi


eDNA_long_all_wide2_sdi$Year -> rownames(eDNA_long_all_wide2_sdi)
eDNA_long_all_wide2_sdi[,-1:-2] -> eDNA_long_all_wide2_sdi

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1 <- eDNA_long_all_wide2_sdi

mds <- vegan::metaMDS(sptran1,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)

```


```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide2_sdi,scor)
score1$rowname <- rownames(score1)
labs$label -> years

as.data.frame(years) -> nmdscov
rownames(nmdscov) <- nmdscov$years
nmdscov$rowname <- rownames(nmdscov)
score2 <- merge(score1, nmdscov, by="rowname", all.x=T, all.y=F)
score2 <- score2[order(score2$year),] 

### from chron cluster
### assign a cluster to each year
score2$cluster <- as.character(labs$clust)


spdf1 %>% 
  dplyr::select(Type,species_color)%>% distinct() %>% 
  arrange(Type) %>%  dplyr::select(species_color) -> color_order_nmds
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path (data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2)) +
  geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= years, colour=cluster), size=10) + 
  scale_colour_manual(values = c("navy","darkred",moon3,color_order_nmds$species_color)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S9_SD_inshore_NMDS.png"),
       width=30,
       height = 30,
       dpi = 400,
      units = c("in"))
```

##### CAP ordination
<br />

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide_sdi ~ two_month_sst + Year , data=sampledf_year_sdi, dist="bray")
summary(vare.cap) -> cap_sum

#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide_sdi

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  top_n(5, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  mutate(., Station ="San Diego Inshore",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_sdi) %>% 
  mutate(.,          `SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2))  + 
  geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
  scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2)) +xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) 

ggsave(filename = here("analysis","figures","CAP_ord_species_sdi.png"),
       width=24,
       height = 12,
       dpi = 400,
      units = c("in"))

```
##### CAP Environmental Variables

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide_sdi ~ two_month_sst + Year , data=sampledf_year_sdi, dist="bray")
summary(vare.cap) -> cap_sum

 
#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide_sdi

as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  mutate(., new_name= case_when(sample=="Year" ~"Year",
                                sample=="two_month_sst" ~"SST ˚C"))-> cap_vars

cap_vars <- as.data.frame(cap_vars)
rownames(cap_vars) <- cap_vars$new_name

# Now add the environmental variables as arrows
arrowmat <- cap_vars


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
 mutate(., Station ="San Diego Offshore",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_sdi) %>% 
  mutate(.,`SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2)) +   geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
   scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))


ggsave(filename = here("analysis","figures","CAP_ord_sdi_env.png"),
       width=16,
       height = 8,
       dpi = 400,
      units = c("in"))

```


# Pt. Conception NMDS & Heatmap
##### Chronological Clustering

<br />
```{r}
b_ID_main_out_site_f %>%   
  filter(., station =="Pt. Conception") %>% 
  dplyr::select(Year, ID_main,Normalized.biomass=  mean) %>% 
  filter(., Normalized.biomass >0) %>% 
  ungroup() %>% 
  group_by(ID_main) %>% 
  dplyr::count() %>% 
  filter(., n> 5) -> species_to_keep_p.c

b_ID_main_out_site_f %>% 
    filter(., station =="Pt. Conception") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sdi$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide_p.c

eDNA_long_all_wide_p.c %>% dplyr::select(Year,station) %>% as.data.frame()-> sampledf_year_ptc
eDNA_long_all_wide_p.c$Year -> rownames(sampledf_year_ptc)

sampledf_year_ptc %>% 
  mutate(., station2 = case_when(station =="Pt. Conception"~"80_60",
                                station =="San Nicholas Island"~"86.7_50",
                                station =="San Diego Offshore"~"93.3_60",
                                station =="San Diego Inshore"~"93.3_30"),
         Year=as.integer(Year)) %>%  
  left_join(two_month_sst, by=c("Year"="Year","station2"="station")) -> sampledf_year_ptc

eDNA_long_all_wide_p.c$Year -> rownames(eDNA_long_all_wide_p.c)
eDNA_long_all_wide_p.c[,-1:-2] -> eDNA_long_all_wide_p.c

#Create a bray-curtis distance matrix
diss_p.c <- vegdist(eDNA_long_all_wide_p.c, method='bray')

#Run the chronological cluster; chclust is the main function from rioja
clust_p.c <- chclust(diss_p.c)


```

```{r}
##Look at the number of clusters.  Then make a category for each cluster
ddata1 <- dendro_data_k(clust_p.c,8)

labs <- ggdendro::label(ddata1)

moon4 <- c("black","navy",moon3,"darkred")
p1 <- plot_ggdendro(ddata1,
                    direction   = "tb",
                    label.size  = 2.5,
                    scale.color = moon4,
                    branch.size = 0.5,
                    expand.y    = 0.2)

p1 <- p1 + theme_void() + expand_limits(x = c(-1, 32))
p1

```
```{r, results='hide'}

#Recreate above chronological clustering dendrogram but using same format as species clustering for figure

BPA_9 <- data.frame(eDNA_long_all_wide_p.c)

dend2 <- BPA_9 %>% vegdist(., method='bray') %>% chclust() %>% as.dendrogram %>% 
  color_branches(k=6, col = moon3)

```

```{r}
#Color code names of species based on trait information
as.data.frame(moon3) -> mooner
mooner$clust <- seq(1,6)

labs %>% 
  left_join(mooner) -> labs

```

<br />

##### Species Clustering

<br />
```{r,fig.height=8, fig.width=8}
## Input is the distance matrix (abundance, etc), with years on x and species on y. and species names as the rownames.
BPA <- data.frame(t(eDNA_long_all_wide_p.c))

colnames(BPA) <- rownames(eDNA_long_all_wide_p.c)
#BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram 

BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram %>% sort(type = "nodes")

#Decide on the number of branches for species, k=3
dend <- BPA %>% vegdist(., method='bray')%>% hclust(method = "complete") %>% as.dendrogram %>% sort(type = "nodes") %>% 
  color_branches(k=5, col=dar3)
plot(dend)
```



```{r}
#Color code names of species based on trait information
as.data.frame(rownames(BPA)) -> species_names

species_names %>% 
  left_join(trait_edna, by=c("rownames(BPA)"="ID_main")) -> species_names

```

##### Figure S10 Heatmap

```{r, fig.height=16, fig.width=8}
#Color palette
some_col_func <- colorspace::sequential_hcl(100,palette="BluYl")


BPA2 <- as.matrix(log10(BPA))
BPA2[BPA2 < 0] <- NA



colseperator <- labs %>% 
  group_by(clust) %>% 
  dplyr::summarise(label =max(label)) %>% 
  left_join(labs)

jpeg(file = here("analysis","figures","S10_Pt.Conception_heatmap.png")
     , width=16,
       height = 12,
       res = 400,
      units = c("in"))
gplots::heatmap.2(BPA2, 
                  main = "",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
dev.off()


gplots::heatmap.2(BPA2, 
                  main = "eDNA Pt. Conception",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
```


##### Figure S11 NMDS

```{r}
b_ID_main_out_site_f %>% 
    filter(., station =="Pt. Conception") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sd$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2_p.c


eDNA_long_all_wide2_p.c$Year -> rownames(eDNA_long_all_wide2_p.c)
eDNA_long_all_wide2_p.c[,-1:-2] -> eDNA_long_all_wide2_p.c

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1 <- eDNA_long_all_wide2_p.c

mds <- vegan::metaMDS(sptran1,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)

```


```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide2_p.c,scor)
score1$rowname <- rownames(score1)
labs$label -> years

as.data.frame(years) -> nmdscov
rownames(nmdscov) <- nmdscov$years
nmdscov$rowname <- rownames(nmdscov)
score2 <- merge(score1, nmdscov, by="rowname", all.x=T, all.y=F)
score2 <- score2[order(score2$year),] 

### from chron cluster
### assign a cluster to each year
score2$cluster <- as.character(labs$clust)


spdf1 %>% 
  dplyr::select(Type,species_color)%>% distinct() %>% 
  arrange(Type) %>%  dplyr::select(species_color) -> color_order_nmds
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path (data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2)) +
  geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= years, colour=cluster), size=10) + 
  scale_colour_manual(values = c("navy","darkred",moon3,color_order_nmds$species_color)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S11_Pt.Conception_NMDS.png"),
       width=30,
       height = 30,
       dpi = 400,
      units = c("in"))
```
##### CAP ordination
<br />

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide_p.c ~ two_month_sst + Year , data=sampledf_year_ptc, dist="bray")
summary(vare.cap) -> cap_sum

#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide_p.c

as.data.frame(vegan::scores(vare.cap, display="species")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  top_n(5, dist) -> top_species
top_species <- as.data.frame(top_species)
rownames(top_species) <- top_species$sample

# Now add the environmental variables as arrows
arrowmat <- top_species


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
  mutate(., Station ="Pt. Conception",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_ptc) %>% 
  mutate(.,          `SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2))  + 
  geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
  scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2)) +xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) 

ggsave(filename = here("analysis","figures","CAP_ord_species_ptc.png"),
       width=24,
       height = 12,
       dpi = 400,
      units = c("in"))

```
##### CAP Environmental Variables

```{r, results="hide",warning=FALSE}
#Run CAP analysis to identify predictor species
vare.cap <- capscale(eDNA_long_all_wide_p.c ~ two_month_sst + Year , data=sampledf_year_ptc, dist="bray")
summary(vare.cap) -> cap_sum

 
#Retain species scores
sppscores(vare.cap) <- eDNA_long_all_wide_p.c

as.data.frame(vegan::scores(vare.cap, display="bp")) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble %>% 
 mutate(dist = sqrt((CAP1 - 0)^2 + (CAP2 - 0)^2)) %>% 
  mutate(., new_name= case_when(sample=="Year" ~"Year",
                                sample=="two_month_sst" ~"SST ˚C"))-> cap_vars

cap_vars <- as.data.frame(cap_vars)
rownames(cap_vars) <- cap_vars$new_name

# Now add the environmental variables as arrows
arrowmat <- cap_vars


# Add labels, make a data.frame
arrowdf <- data.frame(labels = rownames(arrowmat), arrowmat)

# Define the arrow aesthetic mapping
arrow_map <- aes(xend = CAP1, 
                 yend = CAP2, 
                 x = 0, 
                 y = 0, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

label_map <- aes(x = .75 * CAP1, 
                 y = 1.3 * CAP2, 
                 shape = NULL, 
                 color = NULL, 
                 label = labels)

arrowhead = arrow(length = unit(0.02, "npc"))

# Plot CAP
as.data.frame(vare.cap$CCA$wa) %>% 
  rownames_to_column(var = "sample") %>% 
  as.tibble() %>% 
 mutate(., Station ="Pt. Conception",
         Year =as.integer(sample)) %>% 
  left_join(sampledf_year_ptc) %>% 
  mutate(.,`SST ˚C` = two_month_sst) %>% 
  ggplot(aes(x=CAP1, y=CAP2)) +   geom_text_repel(aes(label= Year, color=`SST ˚C`), size=6, fontface='italic') + 	
   scale_color_viridis_c() +
  geom_segment(
    mapping = arrow_map, 
    size = .5, 
    data = arrowdf, 
    color = "grey", 
    arrow = arrowhead
  ) +
  geom_text(
    mapping = label_map, 
    size = 4,  
    data = arrowdf, 
    show.legend = FALSE
  ) + xlab(paste0("CAP1 ","(",label_percent()(cap_sum$cont$importance[2,1]),")")) +ylab(paste("CAP2","(",label_percent()(cap_sum$cont$importance[2,2]),")")) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(),  panel.border = element_rect(colour = "black", fill=NA, size=2))


ggsave(filename = here("analysis","figures","CAP_ord_ptc_env.png"),
       width=16,
       height = 8,
       dpi = 400,
      units = c("in"))

```


# San Nicholas Island NMDS & Heatmap
##### Chronological Clustering

<br />
```{r}
b_ID_main_out_site_f %>%   
  filter(., station =="San Nicholas Island") %>% 
  dplyr::select(Year, ID_main,Normalized.biomass=  mean) %>% 
  filter(., Normalized.biomass >0) %>% 
  ungroup() %>% 
  group_by(ID_main) %>% 
  dplyr::count() %>% 
  filter(., n> 5) -> species_to_keep_sni

b_ID_main_out_site_f %>% 
    filter(., station =="San Nicholas Island") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sdi$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide_sni


eDNA_long_all_wide_sni %>% dplyr::select(Year,station) %>% as.data.frame()-> sampledf_year_sni
eDNA_long_all_wide_sni$Year -> rownames(sampledf_year_sni)

sampledf_year_sni %>% 
  mutate(., station2 = case_when(station =="Pt. Conception"~"80_60",
                                station =="San Nicholas Island"~"86.7_50",
                                station =="San Diego Offshore"~"93.3_60",
                                station =="San Diego Inshore"~"93.3_30"),
         Year=as.integer(Year)) %>%  
  left_join(two_month_sst, by=c("Year"="Year","station2"="station")) -> sampledf_year_sni

eDNA_long_all_wide_sni$Year -> rownames(eDNA_long_all_wide_sni)
eDNA_long_all_wide_sni[,-1:-2] -> eDNA_long_all_wide_sni

#Create a bray-curtis distance matrix
diss_sni <- vegdist(eDNA_long_all_wide_sni, method='bray')

#Run the chronological cluster; chclust is the main function from rioja
clust_sni <- chclust(diss_sni)


```

```{r}
##Look at the number of clusters.  Then make a category for each cluster
ddata1 <- dendro_data_k(clust_sni,8)

labs <- ggdendro::label(ddata1)

moon4 <- c("black","navy",moon3,"darkred")
p1 <- plot_ggdendro(ddata1,
                    direction   = "tb",
                    label.size  = 2.5,
                    scale.color = moon4,
                    branch.size = 0.5,
                    expand.y    = 0.2)

p1 <- p1 + theme_void() + expand_limits(x = c(-1, 32))
p1

```
```{r, results='hide'}

#Recreate above chronological clustering dendrogram but using same format as species clustering for figure

BPA_9 <- data.frame(eDNA_long_all_wide_sni)

dend2 <- BPA_9 %>% vegdist(., method='bray') %>% chclust() %>% as.dendrogram %>% 
  color_branches(k=6, col = moon3)

```

```{r}
#Color code names of species based on trait information
as.data.frame(moon3) -> mooner
mooner$clust <- seq(1,6)

labs %>% 
  left_join(mooner) -> labs

```

<br />

##### Species Clustering

<br />
```{r,fig.height=8, fig.width=8}
## Input is the distance matrix (abundance, etc), with years on x and species on y. and species names as the rownames.
BPA <- data.frame(t(eDNA_long_all_wide_sni))

colnames(BPA) <- rownames(eDNA_long_all_wide_sni)
#BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram 

BPA_1 <- BPA %>% vegdist(., method='bray') %>% hclust %>% as.dendrogram %>% sort(type = "nodes")

#Decide on the number of branches for species, k=3
dend <- BPA %>% vegdist(., method='bray')%>% hclust(method = "complete") %>% as.dendrogram %>% sort(type = "nodes") %>% 
  color_branches(k=5, col=dar3)
plot(dend)
```



```{r}
#Color code names of species based on trait information
as.data.frame(rownames(BPA)) -> species_names

species_names %>% 
  left_join(trait_edna, by=c("rownames(BPA)"="ID_main")) -> species_names

```

##### Figure S12 Heatmap

```{r, fig.height=16, fig.width=8}
#Color palette
some_col_func <- colorspace::sequential_hcl(100,palette="BluYl")


BPA2 <- as.matrix(log10(BPA))
BPA2[BPA2 < 0] <- NA



colseperator <- labs %>% 
  group_by(clust) %>% 
  dplyr::summarise(label =max(label)) %>% 
  left_join(labs)

jpeg(file = here("analysis","figures","S12_San_Nicholas_Island_heatmap.png")
     , width=16,
       height = 12,
       res = 400,
      units = c("in"))
gplots::heatmap.2(BPA2, 
                  main = "",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
dev.off()


gplots::heatmap.2(BPA2, 
                  main = "eDNA San Nicholas Island",
                  srtCol = 60,
                  dendrogram = "both",
                  #density.info="none",
                  Rowv = dend,
                  Colv = dend2, # this to make sure the columns are not ordered
                  trace="none",          
                  #key.xlab = "",
                  #labCol = "",
                  denscol = "lightseagreen",
                  colRow = species_names$species_color,
                  colCol = labs$moon3,
                  margins = c(3,12),
                  density.info = "density",
                  key= TRUE,
                  na.color = "White",
                  colsep= colseperator$x, 
                  sepcolor = "black",
                  col = some_col_func)
```


##### Figure S13 NMDS

```{r}
b_ID_main_out_site_f %>% 
    filter(., station =="San Nicholas Island") %>% 
  dplyr::select(Year,  ID_main,Normalized.biomass=  mean) %>% 
  filter(., ID_main %in% species_to_keep_sd$ID_main) %>% 
  pivot_wider(names_from=ID_main, values_from=Normalized.biomass) %>% 
  as.data.frame()-> eDNA_long_all_wide2_sni


eDNA_long_all_wide2_sni$Year -> rownames(eDNA_long_all_wide2_sni)
eDNA_long_all_wide2_sni[,-1:-2] -> eDNA_long_all_wide2_sni

#NMDS analysis

#Start with the data frame, not the distance matrix

sptran1 <- eDNA_long_all_wide2_sni

mds <- vegan::metaMDS(sptran1,k=2, trace=FALSE, distance = "bray",trymax=100, autotransform = FALSE)
mds ### if stress is under .20, we are okay; if over, may need the change k to 3; 
#stress = .157 at K3
stressplot(mds)

```


```{r}

#Make the nmds plot with ggplot2
###Pull out species info
mdssp <- mds$species
spdf <- data.frame(mdssp)
spdf %>%  drop_na() ->spdf
spdf$species <- rownames(spdf)
#spdf$species1 <- gsub(" ", ".", spdf$species)
spdf1 <- spdf

### This has the habitat and geographic affinities for each species
spdf1 %>% 
  left_join(trait_edna, by=c("species"="ID_main")) -> spdf1
as.factor(spdf1$Type) ->spdf1$Type


#Pull out site info
scor <- scores(mds, display=c("sites","species"))
score1 <- data.frame(eDNA_long_all_wide2_sni,scor)
score1$rowname <- rownames(score1)
labs$label -> years

as.data.frame(years) -> nmdscov
rownames(nmdscov) <- nmdscov$years
nmdscov$rowname <- rownames(nmdscov)
score2 <- merge(score1, nmdscov, by="rowname", all.x=T, all.y=F)
score2 <- score2[order(score2$year),] 

### from chron cluster
### assign a cluster to each year
score2$cluster <- as.character(labs$clust)


spdf1 %>% 
  dplyr::select(Type,species_color)%>% distinct() %>% 
  arrange(Type) %>%  dplyr::select(species_color) -> color_order_nmds
  
ggplot()+ 
  geom_vline(xintercept=0,colour="grey50") +
  geom_hline(yintercept=0,colour="grey50") +
  geom_path (data=score2, aes(x = NMDS1, y = NMDS2, alpha=0.2)) +
  geom_text_repel(data=spdf1,  aes(x = MDS1, y = MDS2, label= species, color=Type), size=6, fontface='italic') + 	
  geom_text_repel(data=score2, aes(x = NMDS1, y = NMDS2,label= years, colour=cluster), size=10) + 
  scale_colour_manual(values = c("navy","darkred",moon3,color_order_nmds$species_color)) +
  theme_bw() +
  labs(list(title="", x="NMDS1",y="NMDS2"))  + xlab("NMDS1") +ylab("NMDS2")+
  theme(  panel.background = element_rect(fill='white', colour='white'),
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          axis.text.x=element_text(size=14, angle=0), 
          axis.text.y=element_text(size=14, angle=0), 
          axis.title.x=element_text(size=18, angle=0), 
          axis.title.y=element_text(size=18, angle=90),
          strip.background = element_rect(fill='white'), 
          strip.text.x=element_text(size=15),
          legend.text=element_text(size=10),
          legend.title=element_text(size=15),
          axis.line=element_line(colour='black'),
          legend.position="none")	

ggsave(filename = here("analysis","figures","S13_San_Nicholas_Island_NMDS.png"),
       width=30,
       height = 30,å
       dpi = 400,
      units = c("in"))
```







